---
title: "ISOM5610 Project_Team 1"
author: "Team 1"
date: "14 December 2018"
output: html_document
---

# Step0: Reading and Cleaning Data

```{r}
setwd("~/MSBA/ISOM5610/final")
income_raw <- read.csv("adult_income.csv", sep = ",", header = TRUE)
income_raw <- income_raw[-1]
summary(income_raw)
str(income_raw) ## we notice "?" needs to be replaced by NA
income_raw[income_raw == "?"] <- NA

library(naniar)
vis_miss(income_raw)
income <- income_raw[complete.cases(income_raw),]
income[,c(2,3,5:8,12,13)] <-  lapply(income[,c(2,3,5:8,12,13)], factor) # drop empty factor levels
sum(is.na(income)) ## all NAs removed
nrow(income)/nrow(income_raw) ## 92.6% of the orginal data are complete and will be used
rm(income_raw)
# income[,c(1,4,9,10,11)] <-  lapply(income[,c(1,4,9,10,11)], as.numeric)
saveRDS(income, "income0.rds") # backup
```

# Step 1: Explore Data

```{r}
income0 <- income # create a duplication for data exploration
summary(income)
str(income)

library(ggplot2)
library(RColorBrewer)

ds1 <- ggplot(income, aes(x=age)) + 
    geom_density(aes(fill=factor(income)),alpha=0.5) +
    labs(title="Density of Income by Age", 
       y="Density of Income", 
       x="Age (Years)") + 
    scale_fill_manual(name = "income",
                      values = c(brewer.pal(7, "Reds")[5], brewer.pal(7, "Blues")[5]),
                      labels = c("Income <= 50K", "Income > 50K"))
plot(ds1)


unique(income[,c(3,4)]) # check relationship between 'education' and 'education_num'. They are exactly the same, matched on a 1-to-1 basis.
#ds2 <- ggplot(income, aes(x=education_num)) + 
#    geom_density(aes(fill=factor(income)),alpha=0.5) +
#    labs(title="Density of Income by Edu_num", 
#       y="Density of Income", 
#       x="Education (Years)") + 
#    scale_fill_manual(name = "income",
#                      values = c(brewer.pal(7, "Reds")[5], brewer.pal(7, "Blues")[5]),
#                      labels = c("Income <= 50K", "Income > 50K"))
#plot(ds2)

bar2 <- ggplot() + geom_bar(data=income, aes(x=education,fill=factor(income)))+
   labs(title="Distribution of Income among Different Education Levels", 
        y="Count by Different Income", x="Education Levels")+
scale_fill_manual(name = "income",values = c(brewer.pal(7, "Reds")[4],
      brewer.pal(7, "Blues")[5]),labels = c("Income <= 50K", "Income > 50K"))+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))

plot(bar2)

ds3 <- ggplot(income, aes(x=capital_gain)) + 
  geom_density(aes(fill=factor(income)),alpha=0.5) + 
  labs(title="Density of Income by Capital Gain", 
       y="Density of Income", x="Capital Gain") + 
  scale_fill_manual(name = "income", values = c(brewer.pal(7, "Reds")[5], 
                                                brewer.pal(7, "Blues")[5]), 
                    labels = c("Income <= 50K", "Income > 50K"))
plot(ds3)

ds4 <- ggplot(income, aes(x=capital_loss)) + 
    geom_density(aes(fill=factor(income)),alpha=0.5) +
    labs(title="Density of Income by Capital Loss", 
       y="Density of Income", 
       x="Capital Loss") + 
    scale_fill_manual(name = "income",
                      values = c(brewer.pal(7, "Reds")[5], brewer.pal(7, "Blues")[5]),
                      labels = c("Income <= 50K", "Income > 50K"))
plot(ds4)

# try net capital gain (loss)
income$capital_net <- income$capital_gain - income$capital_loss
ds4.1 <- ggplot(income, aes(x=capital_net)) + 
    geom_density(aes(fill=factor(income)),alpha=0.5) +
    labs(title="Density of Income by Net Capital Gain (Loss)", 
       y="Density of Income", 
       x="Net Capital Gain (Loss)") + 
    scale_fill_manual(name = "income",
                      values = c(brewer.pal(7, "Reds")[5], brewer.pal(7, "Blues")[5]),
                      labels = c("Income <= 50K", "Income > 50K"))
plot(ds4.1)

# try absolute capital gain (loss) effect
income$capital_logabseff <- log(income$capital_gain + income$capital_loss + 1)
ds4.2 <- ggplot(income, aes(x=capital_logabseff)) + 
    geom_density(aes(fill=factor(income)),alpha=0.5) +
    labs(title="Density of Income by Log Absolute Capital Gain &  Loss Effect", 
       y="Density of Income", 
       x="Log Absolute Capital Gain & Loss Effect") + 
    scale_fill_manual(name = "income",
                      values = c(brewer.pal(7, "Reds")[5], brewer.pal(7, "Blues")[5]),
                      labels = c("Income <= 50K", "Income > 50K"))
plot(ds4.2)


ds5 <- ggplot(income, aes(x=hr_per_week)) + 
    geom_density(aes(fill=factor(income)),alpha=0.5) +
    labs(title="Density of Income by Working Hours", 
       y="Density of Income", 
       x="Working Hours") + 
    scale_fill_manual(name = "income",
                      values = c(brewer.pal(7, "Reds")[5], brewer.pal(7, "Blues")[5]),
                      labels = c("Income <= 50K", "Income > 50K"))
plot(ds5)


bar1 <- ggplot() + geom_bar(data=income, aes(x=type_employer,fill=factor(income)))+
   labs(title="Distribution of Income among Different Types of Employers", 
        y="Count by Different Income", x="Types of Employers")+
scale_fill_manual(name = "income",values = c(brewer.pal(7, "Reds")[4],
brewer.pal(7, "Blues")[5]),labels = c("Income <= 50K", "Income > 50K"))+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))


plot(bar1)



bar3 <- ggplot() + geom_bar(data=income, aes(x=marital,fill=factor(income)))+
   labs(title="Distribution of Income among Different Marital Statuses", 
        y="Count by Different Income", x="Marital Statuses")+
scale_fill_manual(name = "income",values = c(brewer.pal(7, "Reds")[4],
     brewer.pal(7, "Blues")[5]),labels = c("Income <= 50K", "Income > 50K"))+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))


plot(bar3)

bar4 <- ggplot() + geom_bar(data=income, aes(x=occupation,fill=factor(income)))+
   labs(title="Distribution of Income among Different Types of Occupations",
        y="Count by Different Income", x="Types of Occupations")+
scale_fill_manual(name = "income",values = c(brewer.pal(7, "Reds")[4],
      brewer.pal(7, "Blues")[5]), labels = c("Income <= 50K", "Income > 50K"))+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))



plot(bar4)

bar5 <- ggplot() + geom_bar(data=income, aes(x=race,fill=factor(income)))+
   labs(title="Distribution of Income among Different Races", 
        y="Count by Different Income", x="Races")+
scale_fill_manual(name = "income",values = c(brewer.pal(7, "Reds")[4],
      brewer.pal(7, "Blues")[5]), labels = c("Income <= 50K", "Income > 50K"))+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))



plot(bar5)

bar6 <- ggplot() + geom_bar(data=income, aes(x=sex,fill=factor(income)))+
   labs(title="Distribution of Income among Genders", 
        y="Count by Different Income", x="Gender")+
scale_fill_manual(name = "income",values = c(brewer.pal(7, "Reds")[4],
       brewer.pal(7, "Blues")[5]), labels = c("Income <= 50K", "Income > 50K"))



plot(bar6)
```

```{r}
# geographical plot of percentage of income > 50K

binom_income <- function(x) {x==">50K"}
income$income_large <- binom_income(income$income)
library(dplyr)
income_by_country <- tapply(income$income_large, income$country, sum)
count_by_country <- summary(income$country)
countryName <- names(count_by_country)
rindex <- c(4,6,8,9,15,17,33,35,38,39,41)
rrep <- c("Colombia", "Dominican Republic", "El Salvador", "UK", 
          "Netherlands", "Hong Kong",  "Puerto Rico",  
          "South Korea", "Trinidad", "USA", "Bosnia and Herzegovina")
countryName <- replace(countryName, rindex, rrep)


country_table <- data.frame(countryName = countryName,
                           countryCount = count_by_country,
                           countryIncome = income_by_country,
                           countryIncomePct = income_by_country/count_by_country*100)

library(maps)
world_map <- map_data("world")
income_map <- merge(world_map, country_table, by.x = "region", by.y = "countryName", all.x = TRUE, all.y = FALSE)
income_map <- arrange(income_map, group, order)
ggplot(income_map, aes(x = long, y = lat, group = group, fill = countryIncomePct)) + 
  geom_polygon(colour = "grey70") +
  labs(title="Percentage of Income>50K by Country", fill = "Percentage of\nIncome > 50K\n(%)") +
  scale_fill_viridis_c() +
  theme_void()
```

# Step 2: First Recategorization based on data exploration


```{r}
income_rec <- subset(income0,select=-education_num)

## Recategorize the country:
region_table <- data.frame(country=c("Cambodia","Canada","China","Columbia","Cuba",
                  "Dominican-Republic","Ecuador","El-Salvador","England","France","Germany",
                  "Greece","Guatemala","Haiti","Holand-Netherlands","Honduras","Hong","Hungary",
                  "India","Iran","Ireland","Italy","Jamaica","Japan","Laos","Mexico",
                  "Nicaragua","Outlying-US(Guam-USVI-etc)","Peru","Philippines","Poland",
                  "Portugal","Puerto-Rico","Scotland","South","Taiwan","Thailand",
                  "Trinadad&Tobago","United-States","Vietnam","Yugoslavia"),
                  region=c("Asia","NorthAmericas","Asia","Latin America","Latin America",
                    "Latin America","Latin America","Latin America","Europe","Europe","Europe",
                    "Europe","Latin America","Latin America","Europe","Latin America","Asia",
                    "Europe","Asia","Asia","Europe","Europe","Latin America","Asia","Asia",
                    "Latin America","Latin America","NorthAmericas","Latin America","Asia",
                    "Europe","Europe","Latin America","Europe","Asia","Asia","Asia",
                    "Latin America","NorthAmericas","Asia","Europe"))
income_rec <- merge(income_rec, region_table, by.x="country", by.y="country", all.x=TRUE)
income_rec <- subset(income_rec,select=-country)
# summary(income_rec$region)
# 
# table(income_rec$region,income_rec$income)



```
## Data Partition

# Step 3: Fit the Full Model


```{r}
## there are 30,162 observatoins. Ramdomly partition ~80% and ~20% of the data into training and testing set
library(caTools)
set.seed(1111)
spl <- sample.split(income_rec$income, SplitRatio= 0.8)
training_set <- subset(income_rec, spl==TRUE)
testing_set <- subset(income_rec, spl==FALSE)
fit.full <- glm(income~., family=binomial, data=training_set)  ## notice there is perfect separation
summary(fit.full)
```


# Step 4&5 : Detect Perfect Separation and Recategorize again

```{r}

## without pay in type_employer !!
table(income$type_employer,income$income)
income_rec$type_employer[income_rec$type_employer=="Without-pay"] <- "Self-emp-not-inc"
## Preschool in education !!
table(income$education,income$income)
income_rec$education<-as.character(income_rec$education)
income_rec$education[income_rec$education=="Preschool"] <- "Preschool-4th"
income_rec$education[income_rec$education=="1st-4th"] <- "Preschool-4th"
income_rec$education<-as.factor(income_rec$education)

# ## Age !!
# table(income_rec$age,income_rec$income)
# wms <- weighted.mean(c(17,18,19,20,21),c(328,447,594,629,621))
# wml <- weighted.mean(c(81,82,83,84,85,86,88,90),c(13,7,5,8,3,1,3,35))
# income_rec$age[income_rec$age < 22] <-wms
# income_rec$age[income_rec$age > 80] <-wml
# ## education_num !!
# table(income_rec$education_num,income_rec$income)
# wms <- weighted.mean(c(1,2),c(45,151))
# income_rec$education_num[income_rec$education_num<3]<-wms
# ## hr_per_week !!
# 
# table(income_rec$hr_per_week,income_rec$income)

```

# Step 6: Fit full model again

```{r}
set.seed(1111)
spl <- sample.split(income_rec$income, SplitRatio= 0.8)
training_set <- subset(income_rec, spl==TRUE)
testing_set <- subset(income_rec, spl==FALSE)
fit.full <- glm(income~., family=binomial, data=training_set) 
summary(fit.full)
```

Capital_gain still cause convergence problem

```{r}
## replace capital_gain and capital_loss with log absolute capital gain (loss) effect
 income_rec$capital_net <- income$capital_net
 income_rec$capital_gain <- NULL
 income_rec$capital_loss <- NULL
```

```{r}
set.seed(111)
spl <- sample.split(income_rec$income, SplitRatio= 0.8)
training_set <- subset(income_rec, spl==TRUE)
testing_set <- subset(income_rec, spl==FALSE)
fit.full <- glm(income~., family=binomial, data=training_set) 
summary(fit.full)
```

Step7: Stepwise Selection

```{r}
fit.step <- step(fit.full)
summary(fit.step)
# the stepwise result suggests all 8 predictors are useful. We use fit.full as the base model going forward.
fit <- fit.step
```

## capital is useless now!!

Step7.5 : Further Improvement & Interaction


```{r}
## Recategorize the country:
marry_table <- data.frame(marital=c("Divorced","Married-AF-spouse","Married-civ-spouse",
                                    "Married-spouse-absent","Never-married","Separated",
                                    "Widowed"),marital_new=c("Alone","Married",
                                     "Married","Alone","Alone","Alone","Alone"))
income_rec <- merge(income_rec, marry_table, by.x="marital", by.y="marital", all.x=TRUE)
income_rec <- subset(income_rec,select=-marital)
```


```{r}
set.seed(111)
spl <- sample.split(income_rec$income, SplitRatio= 0.8)
training_set <- subset(income_rec, spl==TRUE)
testing_set <- subset(income_rec, spl==FALSE)
fit.full <- glm(income~.-capital_net, family=binomial, data=training_set) 
summary(fit.full)
```
Now all predictors are significant. AUC 87.75%. Check the interaction.


```{r}
attach(income_rec)
r1 <- ggplot() +
  aes(x = type_employer, color = race, group = race, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('type_employer:race')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r1)

r2 <- ggplot() +
  aes(x = type_employer, color = sex, group = sex, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('type_employer:sex')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r2)


r3 <- ggplot() +
  aes(x = type_employer, color = region, group = region, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('type_employer:region')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r3)

r4 <- ggplot() +
  aes(x = type_employer, color = marital_new, group =  marital_new, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('type_employer:marital')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r4)

r5 <- ggplot() +
  aes(x = education, color = race, group = race, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('education:race')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r5)

r6 <- ggplot() +
  aes(x = education, color = sex, group = sex, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('education:sex')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r6)


r7 <- ggplot() +
  aes(x = education, color = region, group = region, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('education:region')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r7)

r8 <- ggplot() +
  aes(x = education, color = marital_new, group =  marital_new, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('education:marital')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r8)

r9 <- ggplot() +
  aes(x = occupation, color = race, group = race, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('occupation:race')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r9)

r10 <- ggplot() +
  aes(x = occupation, color = sex, group = sex, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('occupation:sex')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r10)


r11 <- ggplot() +
  aes(x = occupation, color = region, group = region, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('occupation:region')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r11)

r12 <- ggplot() +
  aes(x = occupation, color = marital_new, group =  marital_new, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('occupation:marital')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r12)

r13 <- ggplot() +
  aes(x = race, color = marital_new, group =  marital_new, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('race:marital')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r13)

r14 <- ggplot() +
  aes(x = sex, color = marital_new, group =  marital_new, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('sex:marital')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r14)

r15 <- ggplot() +
  aes(x = region, color = marital_new, group =  marital_new, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('region:marital')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r15)

r16 <- ggplot() +
  aes(x = region, color = sex, group =  sex, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('region:sex')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r16)

r17 <- ggplot() +
  aes(x = region, color = race, group =  race, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('region:race')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r17)

r18 <- ggplot() +
  aes(x = sex, color = race, group =  race, y = income_rec$income) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line") + ggtitle('sex:race')+
  theme(axis.text.x = element_text(angle = 20, hjust = 1,size=8))
plot(r18)
```

```{r}
library(lmtest)
fit.int1 <- update(fit.full,.~.+type_employer:race)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant but not very good

fit.int1 <- update(fit.full,.~.+type_employer:sex)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant but not very good

fit.int1 <- update(fit.full,.~.+type_employer:marital_new)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant, auc increase to 87.8% !!!!!!!

fit.int1 <- update(fit.full,.~.+education:sex)
lrtest(fit.int1,fit.full)  ## not significant

fit.int1 <- update(fit.full,.~.+education:marital_new)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant but not very good

fit.int1 <- update(fit.full,.~.+occupation:sex)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## not significant 

fit.int1 <- update(fit.full,.~.+occupation:marital_new)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant, auc increase to 87.9% !!!!!!!

fit.int1 <- update(fit.full,.~.+race:marital_new)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant, not so good

fit.int1 <- update(fit.full,.~.+region:marital_new)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## significant, auc increase to 87.7% !!!!!!!

fit.int1 <- update(fit.full,.~.+region:sex)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## not significant


fit.int1 <- update(fit.full,.~.+region:race)
lrtest(fit.int1,fit.full)
summary(fit.int1)  ## not significant
```
add two interaction:

```{r}
fit.int1 <- update(fit.full,.~.+occupation:marital_new)
lrtest(fit.int1,fit.full)
fit.int2 <- update(fit.full,.~.+occupation:marital_new+type_employer:marital_new)
lrtest(fit.int2,fit.int1)
## this one best, auc 87.91%



fit.int3 <- update(fit.full,.~.+occupation:marital_new+type_employer:marital_new+region:marital_new)
lrtest(fit.int3,fit.int2)


```


# Deviance Goodness of Fit Test

```{r}
1-pchisq(fit$deviance, fit$df.residual)
# p-value~1, i.e. it is a good fit model 
```

# Esimtated Probabilities 

```{r}
prob <- predict(fit.int2, newdata=testing_set, type = 'response')
table(testing_set$income, prob > 0.5)

library(InformationValue)
plotROC((testing_set$income==">50K"), prob)

#Double density plot
p <- predict(fit, type='response')
temp_train <- cbind(training_set, p)
p <- prob
temp_test <- cbind(testing_set, p)
ggplot(temp_train, aes(p, color = as.factor(income))) + 
  geom_density(size = 1) +
  ggtitle("Training Set's estimated probabilities") 
```